---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# getdata

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/getdata)](https://CRAN.R-project.org/package=getdata) [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

The goal of {getdata} is to make the experience of getting location data easier and more consistent across a wide variety of sources. {getdata} started as part of the [overedge](https://elipousson.github.io/overedge/) package along with [{maplayer}](https://elipousson.github.io/maplayer/) and [{sfext}](https://elipousson.github.io/sfext/).

{getdata} is designed to work well with location-specific data packages such as {mapmaryland} and {mapbaltimore} and to support reproducible approaches to map-making and place-based data analysis. The package supports downloading access from the following sources

-   ArcGIS FeatureServer and MapServer layers (using {esri2sf})

-   U.S. Census Bureau data (using {tigris})

-   OpenStreetMap (using {osmdata})

-   Socrata Open Data resources (using {RSocrata})

-   Google Sheets (using {googlesheets4})

-   Flickr photos (using {FlickrAPI})

-   Airtable bases (using {httr2} and the Airtable API)

-   Wikipedia articles (using {httr2} and the Wikipedia Geosearch API)

-   Other spatial and tabular data sources including Google MyMaps, GitHub gists, Excel sheets, and any data source already supported by [sf::read_sf()](https://r-spatial.github.io/sf/reference/st_read.html) (see [sfext::read_sf_ext()](https://elipousson.github.io/sfext/reference/read_sf_ext.html) for more details)

The advantage of using {getdata} is that it provides a consistent interface for creating a bounding box for spatial filtering or for querying spatial data by name or id. Where possible, a spatial filter is used before importing or downloading data to avoid the need to load large data files when you are only interested in a small area. The package also provides a consistent approach for handling API tokens and keys (inspired by both the tigris and mapboxapi packages) and for caching data locally.

The related {sfext} package allows {getdata} to supports the easy conversion of tabular data into spatial data. For example, if the source data has coordinates, you can convert the data into an sf object. If data has an address column, you can geocode the data using the {tidygeocoder} package. If the data has a location name column, such as "neighborhood", you can join the data to a simple feature object with the related geometry. You also can turn off these options by setting `geometry = FALSE` for most data access functions.

Lastly, the [format_data()](https://elipousson.github.io/getdata/reference/format_data.html) and [format_sf_data()](https://elipousson.github.io/getdata/reference/format_sf_data.html) functions provide convenient options for working with the data after it is downloaded. While advanced R users may prefer a more customizable options, these functions are designed to support the creation of custom data formatting and access functions such as [format_md_crash_data()](https://elipousson.github.io/mapmaryland/reference/format_md_sf.html) and [get_md_crash_data()](https://elipousson.github.io/mapmaryland/reference/get_md_open_data.html).

This package imports rlang for both non-standard evaluation and error handling and relies on dplyr, purrr, and other tidyverse packages. Suggestions for additional data sources to support, new functions, or improvements to existing functions are welcome.

## Installation

You can install the development version of getdata like so:

``` r
pak::pkg_install("elipousson/getdata")
```

## Example

[get_location_data()](https://elipousson.github.io/getdata/reference/get_location_data.html) is a flexible function for reading and subsetting data. In this example, data is a file path but it can also be a URL, the name of a data set in another package, or a sf object.

```{r}
library(getdata)
library(dplyr)

# location is optional
nc <- get_location_data(data = system.file("shape/nc.shp", package = "sf"))
```

You can use [get_location()](https://elipousson.github.io/getdata/reference/) to get a specific location from a larger simple feature collection that includes a specific type of locations, such as counties in North Carolina. The most basic approach is filtering by name or id:

```{r get_location}
# get_location works with a type sf object and name and id values
location <- get_location(type = nc, name = "Warren", name_col = "NAME")
```

You can then access data within or around this specific location. For example, `get_location_data()` can return all counties within a quarter-mile of Warren County.

```{r}
nearby_counties <-
  get_location_data(
  data = nc,
  location = location,
  dist = 0.25,
  unit = "mi"
)

glimpse(nearby_counties)
```

A similar approach works with other sources although some may require an API key to work.

```{r get_open_data, eval = FALSE}
## Get Q2 2020 vehicle crash data for Cecil County, Maryland
  get_open_data(
    source_url = "https://opendata.maryland.gov",
    data = "65du-s3qu",
    where = "(year = '2020') AND (quarter = 'Q2')",
    name_col = "county_desc",
    name = "Cecil",
    token = Sys.getenv("MARYLAND_OPEN_DATA_API_KEY")
  )
```

You must set or provide an API token or key for [get_open_data()](https://elipousson.github.io/getdata/reference/get_open_data.html), [get_airtable_data()](https://elipousson.github.io/getdata/reference/get_airtable_data.html), [get_flickr_photos()](https://elipousson.github.io/getdata/reference/get_flickr_photos.html) to work. [get_ghseet_data()](https://elipousson.github.io/getdata/reference/get_gsheet_data.html) will require user authentication (handled automatically by the googlesheets4 package).
